---
title: "DifferentialAbundance"
output: 
  flexdashboard::flex_dashboard
---

```{r include=FALSE}
library(tidyverse)
library(phyloseq)
library(DESeq2)
library(plotly)
library(DT)
library(pander)
```

```{r include=FALSE}
plan_name <- snakemake@wildcards[['plan']]
plan <- snakemake@config$differential_abundance[[plan_name]]

prefilter_name <- plan$prefilter
prefilter_string <- snakemake@config$prefilters[[prefilter_name]]$prefilter_function
prefilter_description <- snakemake@config$prefilters[[prefilter_name]]$long_name
design_string <- plan$design
attribute <- plan$attribute
group1_plus <- plan$group1_plus
group2_minus <- plan$group2_minus
```

Info
=====================================

- Comparison: `r plan_name`
- Design: `r design_string`
- Comparison attribute: `r attribute`
- **Positive log fold changes** indicate that expression is higher in **`r paste0(attribute, '=', group1_plus)`** than in **`r paste0(attribute, '=', group2_minus)`**
- **Negative log fold changes** indicate that expression is higher in **`r paste0(attribute, '=', group2_minus)`** than in **`r paste0(attribute, '=', group1_plus)`**
- OTU prefilter function (only OTUs passing filter are tested): ``r prefilter_name`` (`r prefilter_description`)
- OTUs are considered significantly differentially abundant when the adjusted p value (`padj`, or equivalently, False Discovery Rate) is less than 0.1. 

Results table
=====================================

```{r include=FALSE}
res <- readRDS(snakemake@input[['res']])
res %>%
  filter(!is.na(padj)) %>%
  filter(padj < 0.1) -> res.de

res.de.up <- res.de %>%
  filter(log2FoldChange > 0)

res.de.down <- res.de %>%
  filter(log2FoldChange < 0)
```

Differentially abundant OTUs.

```{r echo=FALSE,results='asis'}
cat(paste0("\n\nThese tables are saved as `XLSX/DifferentialAbundanceResults/",plan_name,".xlsx`"))
```

### Higher in `r group1_plus`

```{r echo=FALSE}
DT::datatable(res.de.up,rownames=F,height = '100%',extensions = c('Scroller'),
   options = list( scrollY = 200,  scroller = TRUE,  dom = "Bfrtip"))
```

### Higher in `r group2_minus`

```{r echo=FALSE}
DT::datatable(res.de.down,rownames=F,height = '100%',extensions = c('Scroller'),
   options = list( scrollY = 200,  scroller = TRUE,  dom = "Bfrtip"))
```

Plot
=====================================

```{r include=FALSE}
g <- readRDS(snakemake@input[['ggplot']])
```

Differentially abundant OTUs.

```{r echo=FALSE,results='asis'}
cat(paste0("\n\nFigure saved as `figures/DifferentialAbundanceSinglePlot/",plan_name,".pdf`"))
```

```{r echo=FALSE,results='asis'}
cat(paste0("\n\nPlots of individual OTUs: `figures/DifferentialAbundance/",plan_name,".zip`"))
```

```{r echo=FALSE}
ggplotly(g) %>%
  plotly::layout(boxmode = "group")
```

<!--Include software version info-->

```{r include=FALSE}
absolute.path <- file.path(getwd(),'scripts','SoftwareInfo.Rmd')
```

```{r child=absolute.path}
```